---
title: "ST525 HW 6"
output: pdf_document
author: Nora Quick
---

# Question 1

```{r, include=FALSE}
library(survival)
library(ggplot2)
library(dplyr)
```

## Part (a)
```{r}
ratsA <- data.frame(days = c(123, 166, 188, 188, 192, 192, 201, 209, 213, 216, 216, 219, 230, 234, 246, 265, 304, 226, 244, 266), status = rep(c(1, 0), c(17, 3)))
ratsB <- data.frame(days = c(142, 156, 163, 188, 215, 232, 232, 233, 233, 233, 236, 245, 261, 280, 280, 296, 299, 363, 204, 344), status = rep(c(1, 0), c(18, 2)))
```

```{r}
KM.A <- survfit(Surv(days, status) ~ 1, data = ratsA, conf.type="none")
summ.A <- summary(KM.A, times = c(0,ratsA$days))

plot(KM.A, main = 'Group A', xlab = 'Days',  ylab = 'Survival')

#----
KM.B <- survfit(Surv(days, status) ~ 1, data = ratsB, conf.type="none")
summ.B <- summary(KM.B, times = c(0,ratsB$days))

plot(KM.B, main = 'Group B', xlab = 'Days',  ylab = 'Survival')
```

## Part (b)
```{r}
ltime <- log(summ.A$time)[-c(1,21)]
llsurv <- log(-log(summ.A$surv))[-c(1,21)]

wPlot.A <- data.frame(ltime, llsurv)

ggplot(data = wPlot.A) +
  geom_line(aes(x = ltime, y = llsurv)) +
  geom_point(aes(x = ltime, y = llsurv)) +
  labs(title = 'Group A') +
  ylab('log[-log(Survival Probability)]') +
  xlab('log(time)')

#----
ltime <- log(summ.B$time)[-c(1,21)]
llsurv <- log(-log(summ.B$surv))[-c(1,21)]

wPlot.B <- data.frame(ltime, llsurv)

ggplot(data = wPlot.B) +
  geom_line(aes(x = ltime, y = llsurv)) +
  geom_point(aes(x = ltime, y = llsurv)) +
  labs(title = 'Group B') +
  ylab('log[-log(Survival Probability)]') +
  xlab('log(time)')
```

I believe based on the graphs above that the Kaplan-Meier estimators is a good fit for the survival fit of group A but less so for group B.

## Part (c)
```{r}
fitA <- survreg(formula = Surv(days, status) ~ ., data = ratsA, dist = 'weibull')
summary(fitA)

#----
fitB <- survreg(formula = Surv(days, status) ~ ., data = ratsB, dist = 'weibull')
summary(fitB)
```

The parameter estimates for the Weibull distributions indicate significance with both having a p-value of <2e-16. 

## Part (d)
```{r}
shapeA <- 1/fitA$scale
scaleA <- exp(fitA$coefficients)

plot(KM.A, main = 'Group A', xlab = 'days', ylab = 'survival')
points(ratsA$days, pweibull(ratsA$days, shapeA, scaleA, lower.tail=F),type='p') 

#----
shapeB <- 1/fitB$scale   #shape of Weibull
scaleB <- exp(fitB$coefficients) #scale of Weibull

plot(KM.B, main = 'Group B', xlab = 'days', ylab = 'survival')
points(ratsB$days, pweibull(ratsB$days, shapeB, scaleB, lower.tail=F),type='p') 
```

Assessing again if Weibull distirbution is a good fit for the data I would conclude that, yes, it is a good fit for both group A and group B. 

# Question 2

```{r}
smoke <- read.csv('pharmacoSmoking-old.csv')
head(smoke)
```

```{r}
smoke$admitdate <- smoke$admitdate %>%
  sapply(function(x) x[1]) %>% as.Date(format = c("%m/%d/%y")) 

smoke$fdate <- smoke$fdate %>%
  sapply(function(x) x[1]) %>% as.Date(format = c("%m/%d/%y"))

smoke$time <- difftime(smoke$fdate , smoke$admitdate, units = 'days') %>% as.numeric()

head(smoke)
```

```{r}
smoke$time <- smoke[,1] + 0.1
head(smoke)
```

## Part (a)
```{r}
fitSmoke <- survreg(formula = Surv(time, relapse) ~ 
                  gender + age + grp,
                data = smoke, dist = 'weibull')
summary(fitSmoke)
```
 
## Part (b)
The coefficient of age (0.0006) indicates being older is better after ajusting for gender and grp as well. However, this improvement is not statistically significant with a p-value of 0.90.

## Part (c)
The coefficient of gender (-0.014) indicates being a female is better than being a male, however, with a p-value of 0.90 it is not statistically significant. 

## Part (d)
The coeggicient of grp (-0.13) indicates that being in combination is better than being in patch-only, however, with a p-value of 0.25 we can, again, conclude that it is not statistically significant.


# Question 3

The censoring indicator is indeed switched in this dataset, so you will have to create a new variable that converts it to the expected 1=observed, 0=censored format.

```{r}
color <- read.csv('color.csv')
head(color)
```

```{r}
color$status <- ifelse(color$status==1, 0, ifelse(color$status==0, 1, color$status))
head(color)
```

## Part (a)
```{r}
fitColor <- survreg(formula = Surv(time, status) ~ 
                  group + DVAL + FVAL,
                data = color, dist = 'weibull')
summary(fitColor)
```

## Part (b)
Yes, the failure times do depend on the group with a p-value of 2e-16 in favor of the green group. It also appears that FVAL also statistically matters (p-value = 0.006) for faulure times but DVAL does not (p-value = 0.198). 

## Part (c)
Group green has a longer expected survival time. 
FVAL is also significant and has a 0.99 times (exp(-0.008)) shorter survival time. The larger this value is the shorter the survival time would be and the smaller it get the longer the survival time would be. 

## Part (d)
```{r}
fitexColor <- survreg(formula = Surv(time, status) ~ 
                  group + DVAL + FVAL,
                data = color, dist = 'exponential')
summary(fitexColor)
```

## Part (e)
No, there is no statistically significant data to indicate failure times depend on group, DVAL, or FVAL with p-values of 0.33, 0.84, and 0.44. 

## Part (f)
Yes, two of my conclusions differ. In particular, assuming exponenetial instead of Weibell decreases the significance of the covariates. 

## Part (g)
```{r}
2*(-316.8-(-362.4))
```

The outputted p-value is extremely small (2e-16) indicating that there is a difference between the two fits. We can find that the Weibull distribution is the better fit for this data based on the log likelihood values.




