---
title: "ST525 HW 5"
output: pdf_document
author: Nora Quick
---

# Question 1

Part (a)
```{r}
#15
t15 <- matrix(c(0,1,5,4), ncol = 2)

rownames(t15) <- c('Group A','Group B')
colnames(t15) <- c('d','a')

t15 <- as.table(t15)
t15

#18
t18 <- matrix(c(0,1,4,3), ncol = 2)

rownames(t18) <- c('Group A','Group B')
colnames(t18) <- c('d','a')

t18 <- as.table(t18)
t18

#19
t19 <- matrix(c(0,2,3,1), ncol = 2)

rownames(t19) <- c('Group A','Group B')
colnames(t19) <- c('d','a')

t19 <- as.table(t19)
t19

#20
t20 <- matrix(c(0,1,3,0), ncol = 2)

rownames(t20) <- c('Group A','Group B')
colnames(t20) <- c('d','a')

t20 <- as.table(t20)
t20

#23
t23 <- matrix(c(1,0,1,0), ncol = 2)

rownames(t23) <- c('Group A','Group B')
colnames(t23) <- c('d','a')

t23 <- as.table(t23)
t23
```

Part (b)
```{r}
dA <- c(0,0,0,0,1)
eA <- c(1/2, 1/2, 1/2, 3/4, 1/2)
vA <- c(1/4, 1/4, 1/4, 3/8, 1/4) #??? wasn't explained in ANY material for the class???
dA
eA
vA

dA <- sum(dA)
eA <- sum(eA)
vA <- sum(vA)
dA
eA
vA
```

part (c)
```{r}
logA <- ((1 - 2.75)^2) / 1.375
logA
```

part (d)
```{r}
p_val <- 1 - pchisq(2.227273, df=1)
p_val
```

Based on the resulting p-value I would conclude to fail to reject the null hypothesis that the two groups hit remission at the same time/rate.


# Question 2
Part (a)
```{r}
#10
t10 <- matrix(c(0,1,7,5), ncol = 2)

rownames(t10) <- c('Group A','Group B')
colnames(t10) <- c('d','a')

t10 <- as.table(t10)
t10

#14
t14 <- matrix(c(0,1,7,4), ncol = 2)

rownames(t14) <- c('Group A','Group B')
colnames(t14) <- c('d','a')

t14 <- as.table(t14)
t14

#15
t15 <- matrix(c(0,1,7,3), ncol = 2)

rownames(t15) <- c('Group A','Group B')
colnames(t15) <- c('d','a')

t15 <- as.table(t15)
t15

#16
t16 <- matrix(c(1,0,6,3), ncol = 2)

rownames(t16) <- c('Group A','Group B')
colnames(t16) <- c('d','a')

t16 <- as.table(t16)
t16

#18
t18 <- matrix(c(0,1,5,2), ncol = 2)

rownames(t18) <- c('Group A','Group B')
colnames(t18) <- c('d','a')

t18 <- as.table(t18)
t18

#20
t20 <- matrix(c(1,0,2,2), ncol = 2)

rownames(t20) <- c('Group A','Group B')
colnames(t20) <- c('d','a')

t20 <- as.table(t20)
t20

#21
t21 <- matrix(c(0,1,2,0), ncol = 2)

rownames(t21) <- c('Group A','Group B')
colnames(t21) <- c('d','a')

t21 <- as.table(t21)
t21

#28
t28 <- matrix(c(1,0,1,0), ncol = 2)

rownames(t28) <- c('Group A','Group B')
colnames(t28) <- c('d','a')

t28 <- as.table(t28)
t28
```

Part (b)
```{r}
dA <- c(0,0,0,1,0,1,0,1)
eA <- c(7/13, 7/12, 7/11, 6/10, 5/8, 3/5, 1)
vA <- c(7/26, 7/24, 7/22, 6/20, 5/16, 3/10, 0) #??? 
dA
eA
vA

dA <- sum(dA)
eA <- sum(eA)
vA <- sum(vA)
dA
eA
vA
```

part (c)
```{r}
logA <- ((3 - 4.58)^2) / 1.792
logA
```

part (d)
```{r}
p_val <- 1 - pchisq(1.39308, df=1)
p_val
```

Based on the resulting p-value I conclude to fail to reject the null hypothesis that the two drugs have the same results. 


# Question 3
part (a)
```{r}
tongue <- read.csv('tongue.csv')
head(tongue)
```

part (b)
```{r}
library(survival)
library(survminer)

censored <- Surv(tongue$time, tongue$delta)
fit1 <- survfit(Surv(time, delta) ~ type, data = tongue)
ggsurvplot(fit = fit1, data = tongue, conf.int = TRUE)
```

part (c)
The null hypothesis is that there is no difference between the two types of tumor. 
The alternative hypothesis is that there is a difference btween the two types of tumors.
```{r}
logrank<-survdiff(Surv(time, delta) ~ type, data = tongue)
logrank

peto<-survdiff(Surv(time, delta) ~ type, rho=1, data = tongue)
peto
```

part (d)
Based on the output from part (c) I would conclude that, yes, it is consistent with the plot given in part (b). 


# Question 4
part (a)
```{r}
smoke <- read.csv("pharmacoSmoking-old.csv")
head(smoke)
```

```{r}
smoke$admitdate <- smoke$admitdate %>%
  sapply(function(x) x[1]) %>% as.Date(format = c("%m/%d/%y")) 

smoke$fdate <- smoke$fdate %>%
  sapply(function(x) x[1]) %>% as.Date(format = c("%m/%d/%y"))

smoke$time <- difftime(smoke$fdate , smoke$admitdate, units = 'days') %>% as.numeric()

head(smoke)
```

part (b)
```{r}
library(dplyr)

smoke$f1 <- (smoke$gender == 0 & smoke$grp == 1)
smoke$f2 <- (smoke$gender == 0 & smoke$grp == 2)
smoke$m1 <- (smoke$gender == 1 & smoke$grp == 1)
smoke$m2 <- (smoke$gender == 1 & smoke$grp == 2)

smoke <- smoke %>% mutate(newgroup = case_when(f1 == TRUE ~ 'f1', 
                           f2 == TRUE ~ 'f2', 
                           m1 == TRUE ~ 'm1', 
                           m2 == TRUE ~ 'm2',
                           TRUE ~ 'na'))

head(smoke)

#smoke$newgroup[smoke$newgroup == 'f1'] <- 1
#smoke$newgroup[smoke$newgroup == 'f2'] <- 2
#smoke$newgroup[smoke$newgroup == 'm1'] <- 3
#smoke$newgroup[smoke$newgroup == 'm2'] <- 4
#smoke
```


```{r}
#censored <- Surv(smoke$time, smoke$relapse)
# NOT seperating groups?
fit1 <- survfit(Surv(time, relapse) ~ newgroup, data = smoke)
ggsurvplot(fit = fit1, data = smoke, conf.int = TRUE)
```

part (c)
Null hypothesis: There is no difference in time between admission and relapse based on gender and treatement group.
Alternative hypothesis: There is a difference in time between admission and relapse based on gender and treatement group.
```{r}
logrank<-survdiff(Surv(time, relapse) ~ newgroup, data = smoke)
logrank

peto<-survdiff(Surv(time, relapse) ~ newgroup, rho=1, data = smoke)
peto
```
My conclusion based on these results is that there is moderate evidence that there is no significant difference in the groups based on gender and treatment group. 

part (d)
```{r}
pairwise_survdiff(Surv(time, relapse) ~ newgroup, data = smoke, p.adjust.method = "bonferroni" )
```
This output shows that there is no significant difference between the genders and treatment group. In other words, we fail to reject the null hypothesis. 

